---
title: "APLM hw01"
author: "san teng"
date: "2018年10月11日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r , echo=FALSE}
library(ggplot2)
```

####(Programming work) The director of admission of a small college select 120 stu- dents at random from the new freshman class in a study to determine whether a student’s grade point average (GPA) at the end of freshman year (Y) can be pre- dicted from ACT score (X). The data information are shown in ”CH01PR19.txt”. Assume X and Y follows the simple linear regression model with $cov(\varepsilon_i , \varepsilon_j  ) = 0$ if $i \neq j$, and $cov(\varepsilon_i , \varepsilon_j ) = \sigma^2$ if $i = j$.

```{r ,include=FALSE, echo=FALSE}
setwd("C:/Users/Master/Desktop/黃三/應用統計線性模型")
data <- read.table("CH01PR19.txt" ,header = T)
Y <- data$GPA
X <- cbind(1 ,data$Score)
colnames(X) <- c("1","GPA")

beta_hat <- solve( t(X) %*% X) %*% t(X) %*% Y

g <- ggplot(data, aes(x = Score, y = GPA)) + geom_point() 
g + geom_abline(intercept = beta_hat[1] ,slope = beta_hat[2])

residual <- Y - X %*% beta_hat
sum_of_residual <- sum(residual^2)
MSE <- sum_of_residual / (dim(data)[1] - 2)
```

###(a) the least squares estimators $\hat{\beta} = (\beta_0 , \beta_1)$ is: 
```{r (a) LSE}
beta_hat <- as.vector( solve( t(X) %*% X) %*% t(X) %*% Y )
beta_hat
```

###(b) Plot the fitted regression with data:
```{r (b)Plot}
g <- ggplot(data, aes(x = Score, y = GPA)) + geom_point() 
g + geom_abline(intercept = beta_hat[1] ,slope = beta_hat[2])
```

###我認為這條線(fitted line)並沒有對Score和GPA作出明顯的趨勢呈現。
###Score 和 GPA並沒有到很明顯的正比關係。在Score介於25~30間，有一些點是GPA低的點，使得如果不看這些資料點，則有著Score愈高，GPA愈高的趨勢。

###(c) the residuals and sum of residuals is:

```{r}
residual <- Y - X %*% beta_hat
sum_of_residual <- sum(residual^2)
head(residual)
sum_of_residual
```

###(d) Estimate $\sigma^2$ and $\sigma$ by MSE and $\sqrt{MSE}$:

```{r}
MSE <- sum_of_residual / (dim(data)[1] - 2)
MSE
sqrt(MSE)
```



```{r ,include=FALSE, echo=FALSE}
# lm(GPA ~ Score ,data = data)
summary(lm(GPA ~ Score ,data = data))
#summary(residual) 
#sqrt(t( residual - mean(residual) ) %*% ( residual - mean(residual) ) / (dim(data)[1] - 2))
#g + geom_smooth(method = "lm", se = F)
```





